### 2019-03-13 manuscript/discussion.tex
- backup of discussion up to summary, with addition of Paradox paragraph

%B:  at least two surprising outcomes: (1)  D and P on background of C, and (2) SC not as dead of an end as it was in goldberg2010, if P is partitioned out (can be framed as kind of path-dependence, i.e. qIC path different than rho path, or same as (1), so that I and C are less different on the background of D.
%
Although species are composed of vast assemblages of variable traits, many traits are heritable, and they may affect the propensity of species to perish and multiply \citep{lewontin_1970}, examining such macroevolutionary effects of many traits is rare, and remains challenging.
We investigated the influence of two traits, ploidy and breeding system states, in diversification of Solanaceae.
Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are overshadowed by hidden traits, or breeding system when added to models.  
Interestingly, the difference in diversification rates between the diploids and polyploids is negligible on the background of self-compatibility.
Furthermore, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against the diploid genomic background. %R: furthermore <- in the same vein
Motivated by the widespread recent findings of polyploidy in the ancestry of angiosperms, we also examined and found modest support for diploidization in the history of the family, but without significant effects on inferences of diversification dynamics. %B: how about "modest"? %R- I think modest is a good way to express it because it is the best we can do with chromosome numbers.  %B: changed "polyploidy events" (polyploidy is a noun, only. polyploid either noun or adjective).

Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of intricate hypotheses regarding a heterogenous diversification process.
Below, we discuss key findings and highlight the most interesting and paradoxical implications.
We additionally provide an outline for broader application of phylogenetic comparative methods in the context of interactions among traits and diversification process.

\subsection*{Disentangling trait interactions in a diversification context} %B: Emma may not like "effects,"?
% R- changed the section name to be more accurate

@@

Previous work has shown that SI species in Solanaceae are associated with higher diversification rates compared to their SC counterparts \citep{goldberg_2010}. 
When we only considered the role of breeding system in the diversification process, we were able to replicate the slower net diversification for SC  (\cref{figure:netdivall}C).  %B: we only considered <- we inferred only
Incorporating hidden states again resulted in SI associated with greater net diversification rates than SC in both hidden states  (\cref{figure:netdivall}D). 
However, the overall picture is a bit more complex.
Depending on the background hidden state, it is possible that a class of SC species may experience higher diversification rates than SI species.
Moreover, these SC lineages may not constitute a dead end (with negative diversification rates).

@@

%B: TODO: add below to future goals.
%R: and I: polyploid lineage that has re-diploidized may enhance diversification.
%R: By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with comparative phylogenetic models. 
%In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization  might be a potential issue due to ploidy misclassification.

%B: I will yield to %R here, but I still think there is overwhelming support for this statement, and none of the cited papers show convincingly that, for example, diploidization happens within five orders of magnitude compared with polyploidization.
%Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.
Flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}. 
Following polyploidization, it is possible for genome reorganization, downsizing, and loss \citep{dodsworth_2015, zenil_2016, mandakova_2018}. %B: commas, etc.: importantly rm: "has been observed" this is _only_ true on very, very, short time scales
As a consequence, it is possible that nearly all extant species classified as ``diploid" in our analyses, based on cytogenetic data, are possibly secondary diploids, having undergone both polyploidization and re-diploidization.

Unduly ignoring secondary diploidization would clearly underestimate the rate of transitions from polyploids to diploids, and would then likely cause inflated net diversification rates for diploids, because species considered in the diploid state, may have been ancestrally polyploid, instead. %B: added before conflict/edits
We find strong support for including the diploidization rate parameter ($\delta$) in analyses that include only ploidy \cref{table:bayesfactors}. %B <- %Rosana, inspect this please. It's a merge of conflicted lines.
However, we find only modest support for diploidization in the joint model, and conclusions regarding the relative effects of ploidy and breeding system on diversification are robust to its inclusion.
Also, estimating the rate at which it occurs, based on ploidy level classifications, is highly uncertain (see parameter $\delta$ in \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}), and it can dramatically change ancestral reconstructions (\cref{suppfigure:DPnodipasr,suppfigure:DPasr,suppfigure:DPnodipABasr,suppfigure:DPABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPasr,suppfigure:IDCDCPnodipABasr,suppfigure:IDCDCPABasr}).

% ***
%B: FIXME/TODO Then try to square the oval of history of DP and IC. Specifically, it is difficult to know what to make of the delta models. Their inclusion is 'positively' supported, but yields results that are hard to reconcile with other data. This is one of the more interesting tensions of the paper, in the sense that it binds all of the discussion point and something here has to come to a head. 
%Either what we think we know about SI/SC and reversals is wrong, or something about diploidization inference is wrong. Why? It is hard to square that we don't see IP state, almost ever!, but it's supposed to be the ancestral state, and not just in Solanaceae, across eudicots! Doesn't demand a full treatment here, but it's worth pointing out that if the SI/SC inference is wrong, then we must include reversals, qCI, here and elsewhere and it will change all kinds of inferences. Then, we'll have to explain maintenance of trans-generic polymorphism, ha, ha! Otherwise, something is off with the inference delta >> rho in some cases. If so, we'll have to explain paleo-WGD inference, etc. Either way, we fundamentally misunderstand the evolution of one of these two traits.
% ***

%B: pre-conflict: "Genomic evidence" reads as if the genomes support this. But genomes also do not support the statement (e.g., synteny). They are the data used in studies, not the evidence (it comes from modes of analysis). It would be a bit like referring to evidence in this paper as "evolutionary evidence".
A few lines of evidence indicate that polyploidy may have occurred prior to the origin of Solanaceae, rendering all extant `diploids' secondarily derived. %B: "polyploid event" or "polyploidy" works, but not "polyploidy event"; see above.
First, \citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \textit{Solanum lycopersicum}, may have experienced one or more of WGDs.
A subsequent analysis of synteny between grape and \textit{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \textit{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication \citep{tomato2012}. 
The age of the peak of paralog Ks distances, is approximately 71 million years \citep{tomato2012}. 
If this is the case, then all of the genomes may have been subsequently re-diploidized, yielding the common extant base chromosome numbers in this and related families, n=11-12, presently considered to be diploid \citep{robertson_2011}. 
%Not unlikely, because Arabidopsis n=5.

%B: incorporate ancestral states in absence of I/C consideration to "rag" on \delta? Rag on S-alleles for Ks, or just throw up arms?
%Natural and induced tetraploids from a variety of families, including Fabaceae, Onagraceae, Rosaceae, and Solanaceae, almost invariably lose gametophytic SI (reviewed in \citealt{ramsey_1998,stone_2002,mcclure_2009}).
On the other hand, there is little evidence for the occurrence of re-diploidization in the time frame we considered, within the nightshades.
First, the Ks-inferred duplication long pre-dates the origin of the family \citep{sarkinen_2013}. 
The best-supported, most-recent WGD, which may represent a reversion from polyploidy to diploidy, is older than the lineage under consideration in this study.
At best, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD.
% E: But then how is there SI?! %B: Agreed, we'll cover it.
Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species, including all of those in the relatively cytogenetically conserved `x=12' group, which includes tomato, potato, eggplant, pepper, and tobacco.
%B: Once this is re-arranged, and made to fit with the rest of the discussion, add what this has to do with re-diploidization and how to interpret $\delta$.
It is perhaps not surprising that models diploidization in our study recovered wide credible intervals for the diploidization parameter, $\delta$. 
Despite the limitations of assignment of ploidy levels, we found some support for models that include pararmeter $\delta$, meaning that further study may be warranted, especially in other groups.
%Moreover, rare state changes may not be easily recovered. Emma, you had a clever way to summarize this. 
%     meaning of diploidization parameter

% generality of approach
%
% issues
%     irreversibility SI -> SC assumption

@@

% B: STOPPED HERE

\subsection*{Conclusions}
% problems: this system has a big set of advantages - carefully curated (co-)occurrences of two traits, genetic basis, etc. and everything is still contradictory and opaque.

% R-Here I copied the limitations of the models idea but needs to be reorganized to be cohesive % Fix this a lot more please
% This is the idea on how thorough analyses should be done.
Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as disscussed in  \citet{rabosky_2010, beaulieu_2015}),but also, a problem of inadequate  ``*SSE'' model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focus trait first used to model diversification. We presented a roadmap with a series of analyses that can help scientists to identify if an unknown or unobservable trait is worth pursuing. 
In this roadmap, we have proposed fitting HiSSE models that have effectively shown that some binary states associated with diversification under BiSSE analyses are actually not different in net diversification terms, and instead a hidden state is linked to differences in the diversification rate \citep{beaulieu_2016}.  If that is the case on a system, raising the question of the identity of the hidden state should be considered as the next step in the modeling process. When considering  a second candidate trait, one can model the complex interactions via a MuSSE model, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood. 
Since MuSSE might also suffer from hight type I errors considering more heterogeneity using a hidden state model on top of the multistate SSE is necessary to adequately infer the effect of the complex interactions in the speciation and extinction process.

@@

% outline
% - disentangling effects
% - general value of approach
%  		another kind of confounding processes, not only transition and diversification, but "other diversifications" and "other transitions", esp. correlated
%
% - our study key findings and details
% - effects of Ploidy (as %R)
% - trumped by BS
% - no difference on background of SC 
% - discuss data/models and their long/shortcomings: 
%	ClaSSE, will's idea for stochastic mapping
%	irreversibility SI -> SC assumption
% - how to improve / what next
%  (may be useful to study in SC families)
%
% - pathways
% 
%- diploidization
% - rate, effect
% - history in solanaceae
%
% - summary
%
% B: NEW REORG COMMENTS:

% First, include all straightforward Results and related observations
%
% including delta makes for some nonsense: ancestral states are all P with rampant recent transitions to D (e.g. Fig. S4/S8. Not backed by any data, including synteny, S-allele trans-polymorphism, or anecdotally bearing experimental data about rates of spontaneous polyploidization and diploidization \citep{husband_2012}.

>>>>

%B:  at least two surprising outcomes: (1)  D and P on background of C, and (2) SC not as dead of an end as it was in goldberg2010, if P is partitioned out (can be framed as kind of path-dependence, i.e. qIC path different than rho path, or same as (1), so that I and C are less different on the background of D.
%
Although species are composed of vast assemblages of variable traits, many traits are heritable, and they may affect the propensity of species to perish and multiply \citep{lewontin_1970}, examining such macroevolutionary effects of many traits is rare, and remains challenging.
We investigated the influence of two traits, ploidy and breeding system, in diversification of Solanaceae.
Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are overshadowed by hidden traits, or breeding system when added to models.  
Interestingly, the difference in diversification rates between the diploids and polyploids is negligible on the background of self-compatibility.
Furthermore, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against the diploid genomic background. %R: furthermore <- in the same vein
Motivated by the widespread recent findings of polyploidy in the ancestry of angiosperms, we also examined and found modest support for diploidization in the history of the family, but without significant effects on inferences of diversification dynamics. %B: how about "modest"? %R- I think modest is a good way to express it because it is the best we can do with chromosome numbers.  %B: changed "polyploidy events" (polyploidy is a noun, only. polyploid either noun or adjective).
Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of intricate hypotheses regarding a heterogenous diversification process.
Below, we discuss key findings and highlight the most interesting and paradoxical implications.
Below, we provide an outline for broader application of phylogenetic comparative methods in the context of interactions among traits and diversification process.

\subsection*{Disentangling trait interactions in the context of diversification} %RB: merged two titles; good?

@@

Earlier work has shown that SI species in Solanaceae are associated with higher diversification rates compared to their SC counterparts \citep{goldberg_2010}. 
When we only considered the role of breeding system in the diversification process, we were able to replicate the slower net diversification for SC  (\cref{figure:netdivall}C).  %B: we only considered <- we inferred only
Incorporating hidden states again resulted in SI associated with greater net diversification rates than SC in both hidden states  (\cref{figure:netdivall}D). 
However, the overall picture is a bit more complex.
Depending on the background hidden state, it is possible that a class of SC species may experience higher diversification rates than SI species.
Moreover, these SC lineages may not constitute a dead end (with negative diversification rates).

@@

%R: and I: polyploid lineage that has re-diploidized may enhance diversification.
%R: By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with comparative phylogenetic models. 
%In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization  might be a potential issue due to ploidy misclassification.

%B: I will yield to %R here, but I still think there is overwhelming support for this statement, and none of the cited papers show convincingly that, for example, diploidization happens within five orders of magnitude compared with polyploidization.
%Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.

Our exploratory analyses recover some support for inclusion of diploidization in the history of the family, but without significant effects on inferences of diversification dynamics.
The implications of the results of our diploidization analyses, as well as much of the other recent work on WGDs, are interesting because they conflict with other lines of evidence, especially those concerning the observed simple synteny of genomes in Solanaceae, and patterns of evolution as self-incompatibility loci.

Flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}. 
Following polyploidization, it is possible for genome reorganization, downsizing, and loss \citep{dodsworth_2015, zenil_2016, mandakova_2018}. %B: commas, etc.: importantly rm: "has been observed" this is _only_ true on very, very, short time scales
As a consequence, it is possible that nearly all extant species classified as ``diploid" in our analyses, based on cytogenetic data, are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
Unduly ignoring secondary diploidization would necessarily underestimate the rate of transitions from polyploids to diploids.
It would then likely cause inflated net diversification rates for diploids, because species considered diploid may have been ancestrally polyploid, instead. %B: added before conflict/edits
We find strong support for including the diploidization rate parameter ($\delta$) in analyses that include only ploidy \cref{table:bayesfactors}. %B <- %Rosana, inspect this please. It's a merge of conflicted lines.
However, we find only modest support for diploidization in the joint model, and conclusions regarding the relative effects of ploidy and breeding system on diversification are robust to its inclusion.
Furthermore, estimating the rate of diploidization, based on our ploidy level classifications, is highly uncertain (see parameter $\delta$ in \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}), and yields dramatic changes in ancestral reconstructions (\cref{suppfigure:DPnodipasr,suppfigure:DPasr,suppfigure:DPnodipABasr,suppfigure:DPABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPasr,suppfigure:IDCDCPnodipABasr,suppfigure:IDCDCPABasr}).

Some lines of evidence indicate that polyploidy may have occurred prior to the origin of Solanaceae, rendering all extant `diploids' secondarily derived. %B: "polyploid event" or "polyploidy" works, but not "polyploidy event"; see above.
First, \citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \textit{Solanum lycopersicum}, may have experienced one or more of WGDs.
A subsequent analysis of synteny between grape and \textit{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \textit{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication \citep{tomato2012}. 
The age of the peak of paralog Ks distances, is approximately 71$\pm$19 My \citep{tomato2012}. 
If this is the case, then all of the genomes may have been subsequently re-diploidized, yielding the widespread and common chromosome numbers in this and related families, n=11-12, presently considered to be diploid \citep{robertson_2011}.

On the other hand, there is little evidence for the occurrence of re-diploidization within Solanaceae, since the origin of the family---the time frame we considered.
First, the Ks-inferred duplication likely pre-dates the origin of the family (49 My, HPD 46-53 My \citealt{sarkinen_2013}). 
The best-supported, most-recent WGD \citet{tomato2012} is older than the lineage under consideration in this study.
At best then, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD, but likely before the origin of the family.
Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species, including all of those in the relatively cytogenetically conserved `x=12' group, which includes tomato, potato, eggplant, pepper, and tobacco.
Despite the limitations of assignment of ploidy levels, we found some support for models that include parameter $\delta$, meaning that further study may be warranted, especially in other groups.

Regardless of whether genome polypoidization, followed by widespread diploidization, is a dominant mechanism of genome evolution in Solanaceae, it is clear that our present understanding of the evolution of ploidy and breeding systems is incomplete.
An increasingly forceful weight of evidence seems to support WGDs in the ancestry of many angiosperms. %cite
It is then easily inferred that many species have undergone diploidization, alongside genome size and chromosome number reductions. %cite
Inference of such a common genome upheaval in eudicot history seems to clash with the data indicating that a homologous mechanism of SI, which breaks down nearly invariably in natural and induced tetraploids \citep{stone_2002,mcclure_2009}, has been present continually in many families, including Solanaceae \citep{igic_2006}. 
Most problematically in this context, it is unclear how to explain the maintenance of trans-generic polymorphism at the orthologous S-loci in this and other families.
As well, chromosome numbers and genome comparisons within the family (esp. `x=12' clade, containing \textit{Solanum} and \textit{Nicotiana}) reveal strong conservation.
\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca. 10 inversions and 11 translocations (likely underestimated).
Recovery of such simple relationships would require outstanding convergent loss of duplicated segments.
In either case, all of the approaches make ample assumptions, and it seems that at least some of them will necessitate deep revisions.
Our study, for example, made numerous assumptions in coding trait states (e.g. what constitutes a polyploid or diploid, SI or SC species), in models used for analyses (e.g. irreversibility of SI loss underlain by trans-specific orthologous S-loci), and many more, which could each mislead inference. %B: -> %R: Rosana, take a look at this.
% no discussion of magnitude of \delta vs \rho!
% problems: this system has a big set of advantages - carefully curated (co-)occurrences of two traits, genetic basis, etc. and everything is still contradictory and opaque.

@@

Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as disscussed in  \citet{rabosky_2010, beaulieu_2015}), but also, a problem of inadequate  ``*SSE'' model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focus trait first used to model diversification. 
We presented a roadmap with a series of analyses that can identify whether unknown or unobservable traits are worth pursuing. 
In this roadmap, we have proposed fitting HiSSE models that have effectively shown that some binary states associated with diversification under BiSSE analyses are actually not different in net diversification terms, and instead a hidden state is linked to differences in the diversification rate \citep{beaulieu_2016}.  
If that is the case on a system, raising the question of the identity of the hidden state should be considered as the next step in the modeling process. 
When considering  a second candidate trait, one can model the complex interactions via a MuSSE model, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood. 
Since MuSSE might also suffer from hight type I errors considering more heterogeneity using a hidden state model on top of the multistate SSE is necessary to adequately infer the effect of the complex interactions in the speciation and extinction process.

@@

%B: pre-conflict comment: "Genomic evidence" reads as if the "genomes" support this. But "genomic evidence" also does not support the statement (e.g., widespread synteny). Genomes are the data used in studies, not the evidence (it comes from different modes of genomic analysis).

% outline
% - disentangling effects
% - general value of approach
%  		another kind of confounding processes, not only transition and diversification, but "other diversifications" and "other transitions", esp. correlated
%
% - our study key findings and details
% - effects of Ploidy (as %R)
% - trumped by BS
% - no difference on background of SC 
% - discuss data/models and their long/shortcomings: 
%	ClaSSE, will's idea for stochastic mapping
%	irreversibility SI -> SC assumption
% - how to improve / what next
%  (may be useful to study in SC families)
%

% First, include all straightforward Results and related observations
%
